notes

### 2025-06-30 manuscript/reviews/ResponseReview1.Rmd

> The reviews are through and thoughtful, and I will not reiterate all the details here. For my part, I would like your revision to focus on accessibility. Your paper is a bit challenging-- it's long and a little tough to get through. What can make it easier for less familiar readers? Perhaps more of a basic set up earlier? A glossary defining specific terms (e.g., what is an event? how is it defines? censoring? hazards vs. survival curves, conditional probabilities, etc.)? Also, clarify when readers need to shift from frequentist to Bayesian model... and why? The main consideration here is that wish to make the paper much more accessible-- from start to finish-- for readers who are novices and entirely unfamiliar with this area. This can be done any number of ways, and R3 also includes some suggestions for making these reviews. 
I encourage you to have the revised paper read and pre-reviewed by colleagues who are novices in this area with the explicit goal of making the paper accessible to them-- this will help you hit the more for our broad readership.

> If you choose to submit a revision, please include include a letter detailing your point-by-point responses to each reviewer comment and indicating how you changed the manuscript to address them. The revised manuscript may undergo further peer review. We ask that you submit your revision within three months. Please let us know if you will not be able to meet this deadline.

> Sincerely,

Response: Thank you for providing constructive feedback and allowing us to submit a revision of our manuscript (AMPPS-24-0232).

We tried to focus on accessibility in the revision of our Tutorial in several ways. 
First, we replaced Figure 1 with a simpler Figure that directly conveys our message, and deleted Figures 2 and 3.
Second, we shortened the Introduction and Discussion sections.
Third, we added a section in the supplementary material (section A) in which we visualize the different types of time-to-event data that are obtained in typical RT tasks (detection, discrimination, bistable perception).
We also mention the unique contribution of our work, we removed redundancies, and tried to improve the clarity of the paper.
As such, we have substantially revised the manuscript in accordance with feedback from you as well as the reviewers.

@@

Response: On page 7 of the revised ms., we mention that we measure single-event occurrences that are repeatedly measured in each participant.

> Specific comments:

> The authors use several different terms to refer to time-to-event models throughout the manuscript. The title refers to EHA, and the running head refers to Hazard analysis. Survival analysis is also a common term in Psychology (see tutorials by Lougheed et al., 2019, and Stoolmiller & Snyder, 2006). Disambiguating the various terms early in the manuscript and then constantly using one term throughout the manuscript would be helpful.

Response: We disambiguate the various terms in the revised ms. on page 7, and we now only use one term throughout.

@@

Response: We address our choice for discrete-time methods on pages 7-8 of the revised ms.:

"...the definition of hazard and the type of models employed depend on whether one is using
continuous or discrete time units. As a lab, and mainly for practical reasons, we have much
more experience using discrete-time EHA, and that is the approach that we describe and
focus on in this paper. This choice may seem counter-intuitive, given that RT is typically
treated as a continuous variable. However, continuous forms of EHA require much more
data to estimate the continuous-time hazard (rate) function well. Thus, by trading
a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal
for dealing with typical psychological RT data sets for which there are less than ~200 trials
per condition per experiment. Moreover, as indicated by Allison (2010),
learning discrete-time EHA methods first will help in learning continuous-time methods, so
it seems like a good starting point."

> Page 8-9, excellent and accessible explanations of the survival function in the discrete time framework.

Response: Thank you for the positive feedback. 
To make the manuscript more accessible and shorter and to respond to some other comments from the editor and other reviewers, we moved the explanation of the survival function to the Supplementary Material in the revised ms.

@@

# Reviewer: 2

> Comments to the Author
> Please see attached file for comments.

> Reviewer report for AMPPS-24-0232: “Event History Analysis for psychological time-to-
event data: A tutorial in R with examples in Bayesian and frequentist workflows”
Summary: The authors present a tutorial paper on the analysis of discrete time-to-event
outcomes. They motivate their tutorials with a description of psychological experiments
that produce time-to-event data and explain how simply summarizing event-time
outcomes using means results in a loss of potentially important information. They then
explain how more information can be gleaned from event-time data if they are
summarized using event history analysis methods. They focus specifically on the
analysis of discrete time-to-event outcomes. They provide code for conducting the
analyses in the tutorials on their Github. Overall, I like the idea of this paper and
strongly agree that using methods appropriate for analyzing time-to-event data, rather
than simply summarizing event-time outcomes using means, is very important and
currently underused. While I like the premise of the paper and think that such a paper
could be a valuable contribution to the literature, I do think the paper has room for
improvement in a number of areas (see comments below). The writing is approachable
(which is advantageous for a tutorial paper) but some text in the paper was repetitive
and could be a bit more polished; additional editing with a focus on clarity and brevity
would be beneficial.

Response: Thanks for your constructive and encouraging feedback. As we also noted in our response to the editor, we have made substantial revisions to the manuscript that include removing repetitive text, and placing a stronger focus on clarity and brevity in the revised ms.

@@

Response: We discuss the main contributions above and beyond existing works on pages 5-6 of the revised ms. We address the issue of treating time as discrete or continuous on pages 7-8 of the revised ms. We removed redundancies, get to the tutorial section quicker (page 11) and included more precise explanations of technical jargon.

> <b>Code for the tutorials:</b> I went to the github page and found a folder for each of the 4 tutorials, which included sub-folders for data, figures, and tables. It took a minute for me to realize that the actual R code for each tutorial was in the main folder, not within each sub-folder. This may also be confusing to readers. Additionally, each of the 4 tutorials, along with their sub-tutorials are provided in separate .Rmd files. There were several download steps needed to get all the information. To me, this made the user experience tedious, and made the tutorials seem disjointed from one another. The authors might consider a more comprehensive single Rmd file that is neatly organized with an interactive table of contents (in the knitted html file). To me, this would make the experience much more enjoyable and useful for readers. Given that this is a tutorial, it would also be nice if the the RMarkdowns were somewhat more standalone e.g., by including a brief description of what a life table is and how to interpret the plot.

Response: We explain the location of the files and the folder structure in the README file. In addition we make it clear that s single download step is possible by either cloning the project, or downloading the ZIP file.
We also included a brief description of a life table and how to interpret the plot in the R Markdown tutorial files.
At this point, we also prefer a modular approach to the tutorials rather than combining them. 
We appreciate that not everyone might share this preference, but we think there are many benefits to a modular workflow in this instance.

@@

> <b>Clarity and accuracy: </b>
-There are several instances where jargon terms and key concepts pertinent to survival analysis need to be defined and/or described in slightly more detail than what is currently provided. For example, clearer descriptions of censoring and risk sets (plus what it means to be at risk) should be given in the introduction.

Response: We introduce the definition of right-censoring on page 10 of the revised ms. We discuss what it means to be at risk in section A of the Supplemental Material.

> Additionally, many key details, such as the EHA equations are relegated to supplementary materials. In my opinion, any key information that makes the tutorial more accessible and precise should not be in supplementary materials.

Response: We placed the equations for h(t) and ca(t) in Figure 1 of the revised ms. 
Beyond Figure 1, we still choose to keep any further reference of equations to supplementary materials. 
This is because we feel that one barrier to uptake of these methods is an aversion to equations.
And by placing them in supplementary materials we feel that we strike a nice balance between non-technical accessibility and a fullness of mathemathical explanation for key ideas and concepts that remain easily accessible to interested readers.

@@

> It seems that comparing and contrasting Bayesian and frequentist methods may be outside the scope of a tutorial (pg. 7)

Response: We agree and focus now on the Bayesian models in the revised ms.

> As mentioned above, tutorial 4 – while related – could potentially be its own stand-alone tutorial.

Response: We decided to keep tutorial 4, as we feel that planning future studies is an important and difficult step to take in these kinds of studies, which we wanted to keep close to the rest of the material for ease of understanding a reasonably complete workflow.

### 2025-06-30 manuscript/notes/notes.Rmd
update Notes
commments on the response.

@@

@@

@@

I took out the following and changed it to be far shorter. 
If you want to include any of it in supps, then let's discuss it.
Instead, I chose to focus on picking out the benefits of EHA for research in experimental psychology.
That way, we are being explicit about what we are aiming to do in this section.
And we are therefore making it clear that we are NOT providing a detailed description of EHA.
There is simply not enough space.
And we look confused and contradictory because we don't actually do what we say.

@@

@@

>>>>

-> I removed redundant packages.

@@

-> I got it working but packages are (for now) mentioned in text instead of in footnote...

@@

-> Well done!

@@

I took out the following and changed it to be far shorter. 
If you want to include any of it in supps, then let's discuss it.

-> Section A of the revised Suppl. Material now contains the Figure of RT tasks and discussion of the types of time-to-event data obtained in typical RT tasks.

Instead, I chose to focus on picking out the benefits of EHA for research in experimental psychology.
That way, we are being explicit about what we are aiming to do in this section.
And we are therefore making it clear that we are NOT providing a detailed description of EHA.
There is simply not enough space.
And we look confused and contradictory because we don't actually do what we say.

@@

-> TODO.

@@

-> ok.

# June 30th 2025 #

## section 1.1 ##

- re-combine paragraphs 2 and 3. That way, you don't start a paragraph with "For example ..."

## section 2.3 ##

- para 2. 1 sentence is not a paragraph. is this not finished yet? I don't understand hopw that sentence fits.

## response to reviewers ##

- the formatting really helps distinguish between editor/reviewer comments and our responses. Nice one. it is so much easier to read.

- the below quote is not contradictory, but it sure sounds it. Can we rework it? Do you see what I mean? single events that repeat... Maybe just an extra sentence or two to make the relevant distinction clear.

>Response: On page 7 of the revised ms., we mention that we measure single-event occurrences that are repeatedly measured in each participant.

on line 510 of the response, you say:

>Response: We provided a clearer description of the data.

I think it is essential to always provide page, para and/or line numbers for reviewers so that they can easily check things. This same logic should be applied to all of our responses where necessary. Once the page numbers change, it can be a nightmare to find and verufy changes as a reviewer.
